Analyzing MLG distribution

### 2017-08-15 doc/RMD/MLG-distribution.Rmd
Add graphical legend to mlg distribution plot

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nhaplotypes",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper haplotype",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

@@

@@

```{r graph-16-loci, fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1))
(g16o <- cluster_optimal(graph16loc$total))
plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
```

@@

```{r graph-11-loci, fig.width = 7, fig.height = 7}
(g11o <- cluster_optimal(graph11loc$total))
plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
```

>>>>

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
     # title = "Shared MLGs across regions",
      fill = "Number of\nMLGs",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper MLG"#,
      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
    ))
}

@@

## Graph legend

```{r}
g <- graph_from_adjacency_matrix(matrix(0, 2, 2))
V(g)$size <- c(56, 56 - 32)
V(g)$label <- c("Number of MLGs in Region", "Number of private MLGs")
lay <- as.data.frame(matrix(0, 2, 2, dimnames = list(NULL, c("x", "y"))))
glegend <- create_layout(g, "manual", node.positions = as.data.frame(lay)) %>% ggraph() + 
  geom_node_circle(aes(r = size, fill = size)) + 
  coord_fixed() + 
  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/56, end = 1, guide = "none") + 
  ggrepel::geom_text_repel(aes(label = label), 
                           x = c(20, 0) + 25, 
                           y = c(35, 0), 
                           nudge_x = 350, 
                           segment.size = 1, 
                           arrow = arrow(length = unit(0.15, "native")), 
                           family = "Helvetica") + 
  theme_void() +
  # theme(plot.background = element_rect(color = "black", fill = "grey98")) +
  xlim(c(-60, 480))

glegend
```

@@

```{r graph-16-loci, fig.width = 7, fig.height = 6.25}
par(mfrow = c(1, 1))
(g16o <- cluster_optimal(graph16loc$total))
gg16 <- plot_mlg_graph(graph16loc$total, alt_layout) #+ labs(list(subtitle = "(16 loci)"))
gg16
ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = (6.25/7)*88, units = "mm", scale = 2)
```

```{r, fig.width = 7, fig.height = 6.25}
suppressWarnings(
  gg16cp <- cowplot::ggdraw(xlim = c(0, 1), ylim = c(0, 1)) +
  cowplot::draw_plot(gg16, x = 0, y = 0.05) +
  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
  )
gg16cp
cowplot::ggsave(plot = gg16cp, 
                file = file.path(PROJHOME, "results", "figures", "publication", "mlg-16.pdf"),
                width = 88,
                height = (6.25/7)*88,
                units = "mm",
                scale = 2)
```

@@

```{r graph-11-loci, fig.width = 7, fig.height = 6.25}
(g11o <- cluster_optimal(graph11loc$total))
gg11 <- plot_mlg_graph(graph11loc$total, alt_layout)
gg11
ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = (6.25/7)*88, units = "mm", scale = 2)
```

```{r, fig.width = 7, fig.height = 6.25}
suppressWarnings(
  gg11cp <- cowplot::ggdraw(xlim = c(0, 1), ylim = c(0, 1)) +
  cowplot::draw_plot(gg11, x = 0, y = 0.05) +
  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
  )
gg11cp
cowplot::ggsave(plot = gg11cp, 
                file = file.path(PROJHOME, "results", "figures", "publication", "mlg-11.pdf"),
                width = 88,
                height = (6.25/7)*88,
                units = "mm",
                scale = 2)
```
